Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2017-01-04 Documents/M1/dgM1Curvilinear.tex
Added description of slope limiter and smaller edits to section 2

The model given by Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum} is non-relativistic (i.e., it does not include effects due to strong gravitational fields or interactions with a moving fluid), but is expressed in terms of curvilinear coordinates through the spatial metric tensor $\gamma_{ij}$, which gives squared infinitesimal line-element
\begin{equation}
  ds_{\vect{x}}^{2}=\gamma_{ij}\,dx^{i}\,dx^{j},
\end{equation}
where $\gamma_{ij}$ are the spatial components of the coordinate basis metric tensor.  
We will make two simplifying assumptions: the metric tensor is (1) diagonal and (2) time-independent.  
Specifically we assume the following diagonal form
\begin{equation}
  \gamma_{ij}
  =\mbox{diag}\big[\,\gamma_{11},\,\gamma_{22}(x^{1}),\,\gamma_{33}(x^{1},x^{2})\,\big]
  =\mbox{diag}\big[\,1,\,a(x^{1})^{2},\,b(x^{1})^{2}c(x^{2})^{2}\,\big], 
  \label{eq:threeMetric}
\end{equation}
which is sufficiently general to accommodate Cartesian, cylindrical, and spherical polar coordinates (see Table~\ref{tab:metricFunctions} and Appendix~\ref{app:CurvilinearEuler}).  
Moreover, $\gamma$ is the determinant of the metric tensor, and $\sqrt{\gamma}=abc$.  
The inverse of the spatial metric is denoted $\gamma^{ij}$, so that $\gamma^{ik}\gamma_{kj}=\delta^{i}_{~j}$.  
We can use the spatial metric to raise and lower indices on spatial vectors and tensors; e.g., $\cH_{j}=\gamma_{jk}\,\cH^{k}$.

The two-moment system contains the higher order moments $\cK^{ij}$ --- the symmetric stress tensor.  
In order to form a closed system of equations, the components of the stress tensor must be related to the lower order moments through a closure procedure.  
Following \citet{levermore_1984}, we write
\begin{equation}
  \cK^{ij}
  =\f{1}{2}\,\Big[\,(1-\psi)\,\gamma^{ij}+(3\,\psi-1)\,h^{i}\,h^{j}\,\Big]\,\cJ,
\end{equation}
where $\psi$ is the Eddington factor, $h^{i}=\cH^{i}/\cH$, and $\cH=\sqrt{\cH_{i}\cH^{i}}$.  
We also define the flux factor $h=\cH/\cJ$, and use the analytic expression for the Eddington factor \citep{minerbo_1978} given by 
\begin{equation}
  \psi(h)=\f{1}{3}+\f{2}{3}\,\Big(\,h^{2}\,\big(\,3-h+3h^{2}\,\big)/5\,\Big),
\end{equation}
which approximates the low occupancy limit of the maximum entropy closure for Fermi-Dirac particles given by \citet{cernohorskyBludman_1994}.  
Specifically, we have $\psi(0)=1/3$ (diffusion limit) and $\psi(1)=1$ (streaming limit).

@@

\begin{table}
  \begin{center}
  \caption{Metric functions for Cartesian, Cylindrical, and Spherical coordinate systems.\label{tab:metricFunctions}}
  \begin{tabular}{ccccccccccc}
    \midrule
    Coordinates & $x^{1}$ & $x^{2}$ & $x^{3}$ & $a$ & $b$ & $c$ & $\sqrt{\gamma}$
    & $\pderiv{a}{x^{1}}$ & $\pderiv{b}{x^{1}}$ & $\pderiv{c}{x^{2}}$ \\
    \midrule
    \midrule
    Cartesian & $x$ & $y$ & $z$ & 1 & 1 & 1 & 1 & 0 & 0 & 0 \\
    Cylindrical & $R$ & $z$ & $\phi$ & 1 & $R$ & 1 & $R$ & 0 & 1 & 0 \\
    Spherical & $r$ & $\theta$ & $\phi$ & $r$ & $r$ & $\sin\theta$ & $r^{2}\sin\theta$ & 1 & 1 & $\cos\theta$ \\
    \midrule
    \midrule
  \end{tabular}
  \end{center}
\end{table}

We define the geometry sources in Eq.~\eqref{eq:momentEquationMomentum} by
\begin{equation}
  \cG_{j}
  =\f{1}{2}\,\cK^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}.  
\end{equation}
Specifically, using the time-independent metric tensor in Eq.~\eqref{eq:threeMetric}, the components are
\begin{equation}
  \cG_{1}
  =\f{1}{2}\,\Big(\,\cK^{22}\,\pderiv{\gamma_{22}}{x^{1}}+\cK^{33}\,\pderiv{\gamma_{33}}{x^{1}}\,\Big), \quad
  \cG_{2}
  =\f{1}{2}\,\cK^{33}\,\pderiv{\gamma_{33}}{x^{2}}, \quad\text{and}\quad
  \cG_{3}
  =0.
\end{equation}
Then, by defining the vector of evolved moments $\bcM=\big(\,\cJ,\,\cH_{1},\,\cH_{2},\,\cH_{3}\,\big)^{T}$, the vector of geometry sources $\bcG(\bcM)=\big(\,0,\,\cG_{1},\,\cG_{2},\,\cG_{3}\,\big)^{T}$, the vector of collision sources $\bcC(\bcM)=\big(\,\chi\,(\cJ_{0}-\cJ),\,-\kappa\,\cH_{1},\,-\kappa\,\cH_{2},\,-\kappa\,\cH_{3}\,\big)^{T}$, and the flux vectors $\bcF^{i}(\bcM)=\big(\,\cH^{i},\,\cK_{~1}^{i},\,\cK_{~2}^{i},\,\cK_{~3}^{i}\,\big)^{T}$, we write the system of equations in compact form
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\bcM\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\bcF^{i}(\bcM)\,\Big)
  =\bcG(\bcM)+\bcC(\bcM).
  \label{eq:momentEquationsCompact}
\end{equation}
Eq.~\eqref{eq:momentEquationsCompact} forms the basis for developing DG methods for neutrino transport.

@@

Several books and review articles on the DG method are available now available \citep[see, e.g.,][]{cockburnShu_2001,hesthavenWarburton_2008}, and we will not go into too much detail here.  
However, we review some key concepts to introduce notation, and, since multiple `flavors' of the DG method has emerged, we emphasize specific choices in our implementation.

@@

By defining the cell average
\begin{equation}
  \bcM_{\bK}=\f{|\bK|}{V_{\bK}}\sum_{\bk=\vect{1}}^{\bN}w_{\bk}\,\sqrt{\gamma}_{\bk}\,\bcM_{\bk},
\end{equation}
we obtain the discrete equation for the cell average by multiplying Eq.~\eqref{eq:semidiscreteDiscretized} with $w_{\bk}\,\sqrt{\gamma}_{\bk}$ and summing over all $\bk$
\begin{equation}
  \pd{\bcM_{\bK}}{t}
  =-\f{1}{V_{\bK}}\sum_{i=1}^{d}\sum_{\tilde{\bk}_{i}=\vect{1}}^{\bN}w_{\tilde{\bk}_{i}}
  \Big(\,
    \sqrt{\gamma}(x_{\Hgh}^{i},\tilde{x}_{\tilde{\bk}_{i}}^{i})\,\widehat{\bcF}^{i}(x_{\Hgh}^{i},\tilde{x}_{\tilde{\bk}_{i}}^{i})
    -\sqrt{\gamma}(x_{\Low}^{i},\tilde{x}_{\tilde{\bk}_{i}}^{i})\,\widehat{\bcF}^{i}(x_{\Low}^{i},\tilde{x}_{\tilde{\bk}_{i}}^{i})
  \,\Big)\,|\tilde{\bK}^{i}|
  +\bcG_{\bK}+\bcC_{\bK},
  \label{eq:semidiscreteDiscretizedCellAverage}
\end{equation}
where we have used $\sum_{i=1}^{N}\ell_{i}(\eta)=1$ and $\sum_{i=1}^{N}\partial\ell_{i}/\partial\eta=0$.  
In the absence of sources, Eq.~\eqref{eq:semidiscreteDiscretizedCellAverage} is conservative.

We have now defined the spatial discretization of the moment equations.

@@

\section{Slope Limiters}

\subsection{TVD and TVB Limiters}

For slope limiting, we introduce the modal expansion
\begin{equation}
  (\sqrt{\gamma}\bcM)_{\DG}=\sum_{\bi=\vect{1}}^{\bN}(\widetilde{\sqrt{\gamma}\bcM})_{\bi}\,\phi_{\bi}(\bx),
\end{equation}
where $\{\phi_{\bi}\}_{\bi=\vect{1}}^{\bN}$ are elements of an orthogonal basis and $(\widetilde{\sqrt{\gamma}\bcM})_{\bi}$ are the corresponding expansion coefficients.  
(Here, we let the orthogonal basis be given by normalized Legendre polynomials; $\{\phi_{k}(\eta)\}_{k=1}^{4}=\{\,1,\,\eta,\,\eta^{2}-1/12,\,\eta\,(\eta^{2}-3/20)\,\}$.)  
By orthogonality of the basis functions, we have
\begin{equation}
  (\widetilde{\sqrt{\gamma}\bcM})_{\bi}
  =\cT_{\bi}^{-1}\int_{\bK}(\sqrt{\gamma}\bcM)_{\DG}\,\phi_{\bi}(\bx)\,d\bx,
\end{equation}
where
\begin{equation}
  \cT_{\bi}=\int_{\bK}\phi_{\bi}(\bx)\,\phi_{\bi}(\bx)\,d\bx,
\end{equation}
and $\{\cT_{i}^{-1}\}_{i=1}^{4}=\{\,1,\,12,\,180,\,2800\,\}$.  
\begin{equation}
  (\widetilde{\sqrt{\gamma}\bcM})_{|\vect{1}|}
  \to\text{minmodB}
  \big(\,
    (\widetilde{\sqrt{\gamma}\bcM})_{|\vect{1}|},\,
    \beta\,((\widetilde{\sqrt{\gamma}\bcM})_{|\vect{0}|}-(\widetilde{\sqrt{\gamma}\bcM})_{|\vect{0}|}^{-}),\,
    \beta\,((\widetilde{\sqrt{\gamma}\bcM})_{|\vect{0}|}^{+}-(\widetilde{\sqrt{\gamma}\bcM})_{|\vect{0}|})
  \,\big),
\end{equation}
where $\beta\in[1,2]$.

\citet{cockburnShu_2001}
\begin{equation}
  \mbox{minmodB}(a,b,c)
  =\left\{\begin{array}{ll}
  a \quad\text{if}\quad |a|\le M\,\Delta x^{2} \\
  \mbox{minmod}(a,b,c) \quad \text{otherwise}
  \end{array}\right.,
\end{equation}
where the minmod function is given by
\begin{equation}
  \mbox{minmod}(a,b,c)
  =\left\{\begin{array}{ll}
  s\,\min[\,|a|,|b|,|c|\,] \quad\text{if}\quad $s=\mbox{sign}(a)=\mbox{sign}(b)=\mbox{sign}(c)$ \\
  0 \quad \text{otherwise}.
  \end{array}\right..
\end{equation}

\subsection{Positivity Limiters}

>>>>

The model given by Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum} is non-relativistic (i.e., it does not include effects due to strong gravitational fields or interactions with a moving fluid), but is expressed in terms of curvilinear coordinates through the spatial metric tensor $\gamma_{ij}$, which gives squared infinitesimal line-element
\begin{equation}
  ds_{\vect{x}}^{2}=\gamma_{ij}\,dx^{i}\,dx^{j},
\end{equation}
where $\gamma_{ij}$ are the spatial components of the coordinate basis metric tensor.  
We will make two simplifying assumptions: the metric tensor is (1) diagonal and (2) time-independent.  
Specifically we assume the following diagonal form
\begin{equation}
  \gamma_{ij}
  =\mbox{diag}\big[\,\gamma_{11},\,\gamma_{22}(x^{1}),\,\gamma_{33}(x^{1},x^{2})\,\big], 
  \label{eq:threeMetric}
\end{equation}
which is sufficiently general to accommodate Cartesian, cylindrical, and spherical polar coordinates (see Table~\ref{tab:metricFunctions} and Appendix~\ref{app:CurvilinearEuler}).  
Moreover, $\gamma$ is the determinant of the metric tensor, and $\sqrt{\gamma}=\sqrt{\gamma_{11}\gamma_{22}\gamma_{33}}$.  
The inverse of the spatial metric is denoted $\gamma^{ij}$, so that $\gamma^{ik}\gamma_{kj}=\delta^{i}_{~j}$.  
We can use the spatial metric to raise and lower indices on spatial vectors and tensors; e.g., $\cH_{j}=\gamma_{jk}\,\cH^{k}$.

The two-moment system contains the higher order moments $\cK^{ij}$ --- the symmetric stress tensor.  
In order to form a closed system of equations, the components of the stress tensor must be related to the lower order moments through a closure procedure.  
Following \citet{levermore_1984}, we write
\begin{equation}
  \cK^{i}_{~j}
  =\f{1}{2}\,\Big[\,(1-\psi)\,\delta^{i}_{~j}+(3\,\psi-1)\,h^{i}\,h_{j}\,\Big]\,\cJ,
\end{equation}
where $\psi$ is the Eddington factor, $h^{i}=\cH^{i}/\cH$, and $\cH=\sqrt{\cH_{i}\cH^{i}}$.  
We also define the flux factor $h=\cH/\cJ$, and use the analytic expression for the Eddington factor \citep{minerbo_1978} given by 
\begin{equation}
  \psi(h)=\f{1}{3}+\f{2}{3}\,\Big(\,h^{2}\,\big(\,3-h+3h^{2}\,\big)/5\,\Big),
\end{equation}
which approximates the low occupancy limit of the maximum entropy closure for Fermi-Dirac particles given by \citet{cernohorskyBludman_1994}.  
Specifically, we have $\psi(0)=1/3$ (diffusion limit) and $\psi(1)=1$ (streaming limit).

@@

\begin{table}
  \begin{center}
  \caption{Relevant metric functions for Cartesian, Cylindrical, and Spherical coordinate systems.\label{tab:metricFunctions}}
  \begin{tabular}{ccccccccccc}
    \midrule
    Coordinates & $x^{1}$ & $x^{2}$ & $x^{3}$ & $\gamma_{11}$ & $\gamma_{22}$ & $\gamma_{33}$ & $\sqrt{\gamma}$
    & $\f{1}{\gamma_{22}}\pderiv{\gamma_{22}}{x^{1}}$ & $\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{1}}$ & $\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{2}}$ \\
    \midrule
    \midrule
    Cartesian & $x$ & $y$ & $z$ & 1 & 1 & 1 & 1 & 0 & 0 & 0 \\
    Cylindrical & $R$ & $z$ & $\phi$ & 1 & 1 & $R^{2}$ & $R$ & 0 & $2/R$ & 0 \\
    Spherical & $r$ & $\theta$ & $\phi$ & 1 & $r^{2}$ & $r^{2}\sin^{2}\theta$ & $r^{2}\sin\theta$ & $2/r$ & $2/r$ & $2\cot\theta$ \\
    \midrule
    \midrule
  \end{tabular}
  \end{center}
\end{table}

We define the geometry sources in Eq.~\eqref{eq:momentEquationMomentum} by
\begin{equation}
  \cG_{j}
  =\f{1}{2}\,\cK^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}.  
\end{equation}
Specifically, using the time-independent metric tensor in Eq.~\eqref{eq:threeMetric} and writing in terms of $\cK^{i}_{~j}$, the components are
\begin{equation}
  \cG_{1}
  =\f{1}{2}\,\Big(\,\cK^{2}_{~2}\,\f{1}{\gamma_{22}}\pderiv{\gamma_{22}}{x^{1}}+\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{1}}\,\Big), \quad
  \cG_{2}
  =\f{1}{2}\,\cK^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{2}}, \quad\text{and}\quad
  \cG_{3}
  =0.
\end{equation}
Then, by defining the vector of evolved moments $\bcM=\big(\,\cJ,\,\cH_{1},\,\cH_{2},\,\cH_{3}\,\big)^{T}$, the vector of geometry sources $\bcG(\bcM)=\big(\,0,\,\cG_{1},\,\cG_{2},\,\cG_{3}\,\big)^{T}$, the vector of collision sources $\bcC(\bcM)=\big(\,\chi\,(\cJ_{0}-\cJ),\,-\kappa\,\cH_{1},\,-\kappa\,\cH_{2},\,-\kappa\,\cH_{3}\,\big)^{T}$, and the flux vectors $\bcF^{i}(\bcM)=\big(\,\cH^{i},\,\cK_{~1}^{i},\,\cK_{~2}^{i},\,\cK_{~3}^{i}\,\big)^{T}$, we write the system of equations in compact form
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\bcM\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\bcF^{i}(\bcM)\,\Big)
  =\bcG(\bcM)+\bcC(\bcM).
  \label{eq:momentEquationsCompact}
\end{equation}
Eq.~\eqref{eq:momentEquationsCompact} forms the basis for developing DG methods for neutrino transport.

@@

Several books and review articles on the DG method are now available \citep[see, e.g.,][]{cockburnShu_2001,hesthavenWarburton_2008}, and we will not go into too much detail here.  
However, we review some key concepts to introduce notation, and, since multiple `flavors' of the DG method has emerged, we emphasize specific choices in our implementation.

@@

By defining the cell average
\begin{equation}
  \bcM_{\bK}=\f{|\bK|}{V_{\bK}}\sum_{\bk=\vect{1}}^{\bN}w_{\bk}\,\sqrt{\gamma}_{\bk}\,\bcM_{\bk},
  \label{eq:conservedCellAverage}
\end{equation}
we obtain the discrete equation for the cell average by multiplying Eq.~\eqref{eq:semidiscreteDiscretized} with $w_{\bk}\,\sqrt{\gamma}_{\bk}$ and summing over all $\bk$
\begin{equation}
  \pd{\bcM_{\bK}}{t}
  =-\f{1}{V_{\bK}}\sum_{i=1}^{d}\sum_{\tilde{\bk}_{i}=\vect{1}}^{\bN}w_{\tilde{\bk}_{i}}
  \Big(\,
    \sqrt{\gamma}(x_{\Hgh}^{i},\tilde{x}_{\tilde{\bk}_{i}}^{i})\,\widehat{\bcF}^{i}(x_{\Hgh}^{i},\tilde{x}_{\tilde{\bk}_{i}}^{i})
    -\sqrt{\gamma}(x_{\Low}^{i},\tilde{x}_{\tilde{\bk}_{i}}^{i})\,\widehat{\bcF}^{i}(x_{\Low}^{i},\tilde{x}_{\tilde{\bk}_{i}}^{i})
  \,\Big)\,|\tilde{\bK}^{i}|
  +\bcG_{\bK}+\bcC_{\bK},
  \label{eq:semidiscreteDiscretizedCellAverage}
\end{equation}
where we have used $\sum_{i=1}^{N}\ell_{i}(\eta)=1$ and $\sum_{i=1}^{N}\partial\ell_{i}/\partial\eta=0$.  
In the absence of sources, Eq.~\eqref{eq:semidiscreteDiscretizedCellAverage} is conservative.

This defines the spatial discretization of the moment equations.

@@

\section{Polynomial Limiting}

During time integration we apply limiters that modify the polynomial representation in elements so as to prevent development and growth of nonphysical oscillations (slope limiters; Sec.~\ref{sec:slopeLimiters}) or nonphysical states (positivity limiters; Sec.~\ref{sec:positivityLimiters}).

\subsection{Slope Limiters}
\label{sec:slopeLimiters}

For slope limiting, we introduce the modal expansion
\begin{equation}
  \bcM_{\DG}(\bx)=\sum_{\bi=\vect{1}}^{\bN}\bcC_{\bi}\,\phi_{\bi}(\bx),
  \label{eq:conservedModalExpansion}
\end{equation}
where $\{\phi_{\bi}\}_{\bi=\vect{1}}^{\bN}$ are elements of an orthogonal basis and $\bcC_{\bi}$ are the corresponding expansion coefficients (not to be confused with the collision term in Eq.~\eqref{eq:momentEquationsCompact}).  
Here, we let the orthogonal basis be constructed from the tensor product of one-dimensional, normalized Legendre polynomials up to degree 3; i.e., $\{\phi_{k}(\eta)\}_{k=1}^{4}=\{\,1,\,\eta,\,\eta^{2}-1/12,\,\eta\,(\eta^{2}-3/20)\,\}$.  
By orthogonality of the basis functions, the expansion coefficients are easily obtained by projection of the nodal representation in Eq.~\eqref{eq:conservedNodalExpansion} onto the modal basis in Eq.~\eqref{eq:conservedModalExpansion}; i.e.,
\begin{equation}
  \bcC_{\bi}
  =\langle\,\phi_{\bi},\,\phi_{\bi}\,\rangle_{\bK}^{-1}\sum_{\bj=\vect{1}}^{\bN}\langle\,\phi_{\bi},\,\ell_{\bj}\,\rangle_{\bK}\,\bcM_{\bj}
\end{equation}
where
\begin{equation}
  \langle\phi_{\bi},\phi_{\bi}\rangle_{\bK}=\int_{\bK}\phi_{\bi}(\bx)\,\phi_{\bi}(\bx)\,d\bx
  \quad\text{and}\quad
  \langle\phi_{\bi},\ell_{\bj}\rangle_{\bK}=\int_{\bK}\phi_{\bi}(\bx)\,\ell_{\bj}(\bx)\,d\bx.
  \label{eq:projectionIntegrals}
\end{equation}
The integrals in Eq.~\eqref{eq:projectionIntegrals} can be computed exactly with a sufficiently accurate quadrature rule.  
The projection is a linear transformation, and the projection matrix integrals can be computed once and stored at program initialization.  
Also note that the projection matrix does not depend on the metric function, and is therefore the same for all elements, which limits storage requirements.

The elements of the multidimensional modal basis are ordered by increasing polynomial degree; i.e., 
\begin{equation}
  \{\phi_{1}(\eta^{1})\phi_{1}(\eta^{2})\ldots\phi_{1}(\eta^{d}),\,\phi_{2}(\eta^{1})\phi_{1}(\eta^{2})\ldots\phi_{1}(\eta^{d}),\,\phi_{1}(\eta^{1})\phi_{2}(\eta^{2})\ldots\phi_{1}(\eta^{d}),\ldots\},
\end{equation}
so that the first expansion coefficient represents the average value of $\bcM$ in $\bK$, $\bcC_{|\vect{0}|}=\langle\bcM\rangle_{\bK}$ (the coefficient of the basis polynomial with total degree $0$).  
The second expansion coefficient, $\bcC_{|\vect{1}|}^{1}$ (coefficient of the basis function with total degree $1$ varying in the first dimension), represents the first derivative of $\bcM$ in the first dimension, and so on.

The limiting procedure first computes, for each dimension $i=1,\ldots,d$, the limited slopes $\widetilde{\bcC}_{|\vect{1}|}^{i}$ by comparing the variation within $\bK$, $\bcC_{|\vect{1}|}^{i}$, with the variation in the cell average by considering the neighboring elements
\begin{equation}
  \widetilde{\bcC}_{|\vect{1}|}^{i}
  =\text{minmodB}
  \big(\,
    \bcC_{|\vect{1}|}^{i},\,
    \beta\,(\langle\bcM\rangle_{\bK}-\langle\bcM\rangle_{\bK}^{i,-}),\,
    \beta\,(\langle\bcM\rangle_{\bK}^{i,+}-\langle\bcM\rangle_{\bK})
  \,\big),
  \label{eq:limitedSlopes}
\end{equation}
where the limiter parameter $\beta\in[1,2]$ controls how aggressive limiting is applied.  
($\beta=2$ is the least aggressive, and, in many cases, leads to the most accurate results.)  
The $\text{minmodB}$ function is given by \citet{cockburnShu_2001} as
\begin{equation}
  \mbox{minmodB}(a,b,c)
  =\left\{\begin{array}{ll}
  a \quad\text{if}\quad |a|\le M\,\Delta x^{2} \\
  \mbox{minmod}(a,b,c) \quad \text{otherwise},
  \end{array}\right.
  \label{eq:minmodB}
\end{equation}
and the usual minmod function is given by
\begin{equation}
  \mbox{minmod}(a,b,c)
  =\left\{\begin{array}{ll}
  s\,\min[\,|a|,|b|,|c|\,] \quad\text{if}\quad s=\mbox{sign}(a)=\mbox{sign}(b)=\mbox{sign}(c) \\
  0 \quad \text{otherwise}.
  \end{array}\right.
  \label{eq:minmod}
\end{equation}
In Eq.~\eqref{eq:minmodB}, the parameter $M>0$ is proportional to the second derivative, and can be adjusted to prevent clipping at smooth extrema.  
For $M=0$, the $\text{minmodB}$ functions reduces to the usual minmod function in Eq.~\eqref{eq:minmod}.  
In Eq.~\eqref{eq:limitedSlopes}, $\langle\bcM\rangle_{\bK}^{i,-/+}$ is the cell average in the element to the left/right in the $i$-th dimension of the element under consideration.

After computing the limited slopes in Eq.~\eqref{eq:limitedSlopes}, they are compared with the original slopes.  
Then, if
\begin{equation}
  |\widetilde{\bcC}_{|\vect{1}|}^{i}-\bcC_{|\vect{1}|}^{i}|/\max\big(|\widetilde{\bcC}_{|\vect{1}|}^{i}|,|\bcC_{|\vect{1}|}^{i}|\big)>\text{tol},
\end{equation}
where we set $\text{tol}=10^{-2}$, we replace the modal expansion in Eq.~\eqref{eq:conservedModalExpansion} with the limited polynomial
\begin{equation}
  \bcM_{\DG}(\bx)
  \to\widetilde{\bcM}_{\DG}(\bx)
  =\langle\bcM\rangle_{\bK}+\sum_{i=1}^{d}\widetilde{\bcC}_{|\vect{1}|}^{i}\,\phi_{|\vect{1}|}^{i}(\bx).  
  \label{eq:conservedModalExpansionLimited}
\end{equation}
We limit all the fields in $\bcM$ if the difference in any of the coefficients exceeds the tolerance.  
This limiting procedure locally destroys the high-order accuracy of the DG method, but leads to non-oscillatory evolution when sharp features are present in the solution.

The limiting procedure described here operates on the coefficients of the modal expansion of $\bcM$, not that of $\sqrt{\gamma}\bcM$.  
As a consequence, the average $\langle\bcM\rangle_{\bK}$ is preserved in the limiting process, and not the average given in Eq.~\eqref{eq:conservedCellAverage}, which defines the conserved quantities governed by the two-moment model in curvilinear coordinates.  
(We have found that limiting a polynomial expansion of $\sqrt{\gamma}\bcM$ leads to inaccurate results, especially near coordinate singularities.)  
To enforce conservation, we therefore introduce a correction step, similar in spirit to that discussed by \citet{radiceRezzolla_2011}.  
To this end we define $\widetilde{\bcC}=\big(\bcC_{|\vect{0}|},\,\widetilde{\bcC}_{|\vect{1}|}^{1},\ldots,\,\widetilde{\bcC}_{|\vect{1}|}^{d}\big)^{T}$ and $\widehat{\bcC}=\big(\widehat{\bcC}_{|\vect{0}|},\,\widehat{\bcC}_{|\vect{1}|}^{1},\ldots,\,\widehat{\bcC}_{|\vect{1}|}^{d}\big)^{T}$, and find $\widehat{\bcC}$ by solving
\begin{equation}
  \min_{\widehat{\bcC}}\f{1}{2}||\widetilde{\bcC}-\widehat{\bcC}||^{2}
\end{equation}
subject to the constraint
\begin{equation}
  \widehat{\bcM}_{\bK}=\f{1}{V_{\bK}}\int_{\bK}\widehat{\bcM}_{\DG}\,\sqrt{\gamma}\,d\bx\equiv\bcM_{\bK}.  
\end{equation}
This linear equality-constrained least squares problem can be solved, e.g., using LAPACK's DGGLSE subroutine.

\subsection{Positivity Limiters}
\label{sec:positivityLimiters}
